---
# title: "ERGMnetworks"
# author: "Laura Dozal"
# date: "December 14, 2023"
output: html_document
---

## ERGM

*"...how can we know how our observed network density compares to chance?"*

With ERGMs, we ask: given a sample of networks with like characteristics as the observed network (g(y)), what is the likelihood that these simulations produce the observed network? ... we can compare the observed network statistics with statistics from simulated networks with like characteristics.

ERGMs attempt to bring a linear modeling approach to network estimation by assuming that these inherent interdependencies depend mostly on network characteristics that a savvy modeler can explicitly specify. 

**Network density** is the ratio of the number of edges to the maximum possible number of edges


*Quotes from:* https://medium.com/@ben.roston/exponential-random-graph-models-ergms-network-analysis-and-you-2d7d756345e8

https://socialabstractions-blog.tumblr.com/post/53391947460/exponential-random-graph-models-in-python

```{r load_libraries}
library(dplyr)
library(igraph)
# install.packages('network')
library(network)
```

```{r load data}

# load .csv
data <- read.csv("data/keywords.csv")
head(data)
```

```{r}
# select the Source and Destination airport columns
keyword_edgelist <- data %>%
  select(keywords, publisher_categories)

# rename columns to work with igraph functions
colnames(keyword_edgelist) <- c("source", "target")

# load edgelist into an networkobject
keywords <- network(keyword_edgelist, loops = TRUE, multiple = TRUE)
keywords
```

