---
# title: "ERGMnetworks"
# author: "Laura Dozal"
# date: "December 14, 2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) #echo = TRUE, 
```


# ERGMs (Exponential Random Graph Model)


*"...how can we know how our observed network density compares to chance?"*

With ERGMs, we ask: given a sample of networks with like characteristics as the observed network (g(y)), what is the likelihood that these simulations produce the observed network? ... we can compare the observed network statistics with statistics from simulated networks with like characteristics.

ERGMs attempt to bring a linear modeling approach to network estimation by assuming that these inherent interdependencies depend mostly on network characteristics that a savvy modeler can explicitly specify. 

**Network density** is the ratio of the number of edges to the maximum possible number of edges


*Quotes from:* https://medium.com/@ben.roston/exponential-random-graph-models-ergms-network-analysis-and-you-2d7d756345e8

https://socialabstractions-blog.tumblr.com/post/53391947460/exponential-random-graph-models-in-python


```{r load libraries, echo = FALSE, warning=FALSE, message = FALSE}

library(dplyr)
library(igraph)
# install.packages('network')
# install.packages('ergm', version = ">= 1.1.1")
# install.packages('statnet')
library(ergm)
library(statnet)
library(network)
```

```{r load data, echo = FALSE}

# remove.packages("rlang")
# install.packages("rlang")

# remove.packages("vctrs")
# install.packages("vctrs")

# remove.packages("digest")
# install.packages("digest")
# remove.packages("rmarkdown")
# install.packages("rmarkdown")

# load .csv
data <- read.csv("data/keywords.csv")
authors <- read.csv("data/authors.csv")
# head(data)
```

## Keywords
```{r create network, echo=FALSE}
# select the Source and Destination airport columns
keyword_edgelist <- data %>%
  select(keywords, publisher_categories)
# print(head(keyword_edgelist))

# rename columns to work with igraph functions
colnames(keyword_edgelist) <- c("source", "target")
# print(keyword_edgelist)
# load edgelist into an networkobject
kwd <- network(keyword_edgelist, loops = FALSE, multiple = TRUE)
# summary(kwd)
# class(kwd)
```


```{r review network density, echo=FALSE, warning=FALSE, include=FALSE}

require(statnet)
kwd_graph <- graph(cbind(keyword_edgelist$source, keyword_edgelist$target))
print(kwd_graph)
kwd_gMemb <- kwd_graph$membership
kwd_gCsize <- kwd_graph$csize
kwd_gCount <- kwd_graph$no
# statnet::gden(kwd_gCsize)
# components(kwd)

# centralization.betweenness(kwd_graph)
# centralization.closeness(kwd_graph)
# centralization.degree(kwd_graph)

```

### Review Network Connections
Let's take a look at our keyword network which has 273 edges, no pairs nor triads.
```{r edges, echo=FALSE, warning=FALSE, message = FALSE}

#how many edges does our network have?
summary(kwd ~ edges)
#how many paired connections does the network have?
summary(kwd ~ cycle(k = 2))
#how many triad connections does the network have?
summary(kwd ~ cycle(k = 3))


```

### Apply ERGM Model
Here our Maximum Likelihood for edges is -5.01 which is essentially our slope, meaning the frequency of keyword connections decrease. We can see this with the MCMC simulations in the next chunck.
```{r ergm model, echo=FALSE, warning=FALSE, message = FALSE}

#create our ERGM model using edges
# set a seed to control for ranomization
kwd_ergm <- ergm(kwd ~ edges,
               control=control.ergm(seed=40))
# review the Maximum likelihood
kwd_ergm

```

### Simulate networks to validate the ERGM Model
Here we pull a simulated network to show predicted attributes for #20. Plots for  Random Simulated Network #20, #53, and #80 are also below.
```{r ergm simulations, echo=FALSE, warning=FALSE, message = FALSE}


#now create simulations of our model to validate our findings
# Bootstrapping - here we simulate 100 networks
kwd_ergm_sims <- simulate(
  kwd_ergm, 
  nsim = 100,
  control = control.simulate.ergm(
    MCMC.burnin = 5000000,
    MCMC.interval = 25000
  ),
  verbose = FALSE
  
)
#select a random simulated network to check its attributes
kwd_ergm_sims[[20]]
plot(kwd_ergm_sims[[20]], main = "Random Simulated Network #20")
plot(kwd_ergm_sims[[20]], main = "Random Simulated Network #53")
plot(kwd_ergm_sims[[87]], main = "Random Simulated Network #87")

```


### Histograms of Connections in our 100 Similated Networks

Below we see the number of pairs and triads from our similated networks for our keywords has a higher frequency of zero or one connection. This coincides with our own network outcomes and so we can validate that our keyword network is a common phenomena. 
```{r plot simulations, echo=FALSE, warning=FALSE, message = FALSE}

# class(kwd_ergm_sims)
simulated_n_pair <- sapply(
  kwd_ergm_sims, 
  function(sim_network) summary(sim_network ~ cycle(2))
)

simulated_n_pair <- as.data.frame(simulated_n_pair)
table_pairs <- table(simulated_n_pair$simulated_n_pair)

barplot(table_pairs,
        border="#69b3a2", col="white",
        xlab = 'Paired Connections in 100 Network Simulations',
        ylab = 'Count',
        ylim = c(0, 45))

simulated_n_triads <- sapply(
  kwd_ergm_sims, 
  function(sim_network) summary(sim_network ~ cycle(3))
)
simulated_n_triads <- as.data.frame(simulated_n_triads)
table_triads <- table(simulated_n_triads$simulated_n_triads)
barplot(table_triads,
        border="#69b3a2", col="white",
        xlab = 'Tryad Connections in 100 Network Simulations',
        ylab = 'Count',
        ylim = c(0, 45))


```


## Authors
```{r create author network, echo=FALSE}
# select the Source and Destination airport columns
authors_edgelist <- authors %>%
  select(researcher, publisher_categories)
# print(head(authors_edgelist))

# rename columns to work with igraph functions
colnames(authors_edgelist) <- c("source", "target")
# print(authors_edgelist)
# load edgelist into an networkobject
researcher <- network(authors_edgelist, loops = FALSE, multiple = TRUE)
# summary(researcher)
# class(researcher)
```


```{r review authro network density, echo=FALSE, warning=FALSE, include=FALSE}

require(statnet)
researcher_graph <- graph(cbind(authors_edgelist$source, authors_edgelist$target))
print(researcher_graph)
researcher_gMemb <- researcher_graph$membership
researcher_gCsize <- researcher_graph$csize
researcher_gCount <- researcher_graph$no
# statnet::gden(kwd_gCsize)
components(researcher)

# centralization.betweenness(kwd_graph)
# centralization.closeness(kwd_graph)
# centralization.degree(kwd_graph)

```

### Review Network Connections
Let's take a look at our researcher network which has 273 edges, no pairs nor triads.
```{r author edges, echo=FALSE, warning=FALSE, message = FALSE}

#how many edges does our network have?
summary(researcher ~ edges)
#how many paired connections does the network have?
summary(researcher ~ cycle(k = 2))
#how many triad connections does the network have?
summary(researcher ~ cycle(k = 3))


```

### Apply ERGM Model
Here our Maximum Likelihood for edges is -8.77 which is essentially our slope, meaning the frequency of researcher connections decrease. We can see this with the MCMC simulations in the next chunck.
```{r authro ergm model, echo=FALSE, warning=FALSE, message = FALSE}

#create our ERGM model using edges
# set a seed to control for ranomization
researcher_ergm <- ergm(researcher ~ edges,
               control=control.ergm(seed=40))
# review the Maximum likelihood
researcher_ergm

```

### Simulate networks to validate the ERGM Model
Here we pull a simulated network to show predicted attributes for #20. Plots for  Random Simulated Network #20, #53, and #80 are also below.
```{r author ergm simulations, echo=FALSE, warning=FALSE, message = FALSE}


#now create simulations of our model to validate our findings
# Bootstrapping - here we simulate 100 networks
researcher_ergm_sims <- simulate(
  researcher_ergm, 
  nsim = 100,
  control = control.simulate.ergm(
    MCMC.burnin = 5000000,
    MCMC.interval = 25000
  ),
  verbose = FALSE
  
)
#select a random simulated network to check its attributes
researcher_ergm_sims[[20]]
plot(researcher_ergm_sims[[20]], main = "Random Simulated Network #20")
plot(researcher_ergm_sims[[20]], main = "Random Simulated Network #53")
plot(researcher_ergm_sims[[87]], main = "Random Simulated Network #87")

```


### Histograms of Connections in our 100 Similated Networks

Below we see the number of pairs and triads from our similated networks for our Researchers has a higher frequency of zero or one connection. This coincides with our own network outcomes and so we can validate that our keyword network is a common phenomena. 
```{r plot author simulations, echo=FALSE, warning=FALSE, message = FALSE}

# class(kwd_ergm_sims)
simulated_n_pair <- sapply(
  researcher_ergm_sims, 
  function(sim_network) summary(sim_network ~ cycle(2))
)

simulated_n_pair <- as.data.frame(simulated_n_pair)
table_pairs <- table(simulated_n_pair$simulated_n_pair)

barplot(table_pairs,
        border="#69b3a2", col="white",
        xlab = 'Paired Connections in 100 Network Simulations',
        ylab = 'Count',
        ylim = c(0, 85))

simulated_n_triads <- sapply(
  researcher_ergm_sims, 
  function(sim_network) summary(sim_network ~ cycle(3))
# )
simulated_n_triads <- as.data.frame(simulated_n_triads)
table_triads <- table(simulated_n_triads$simulated_n_triads)
barplot(table_triads,
        border="#69b3a2", col="white",
        xlab = 'Tryad Connections in 100 Network Simulations',
        ylab = 'Count',
        ylim = c(0, 85))


```

